---
title: "HW 1"
author: "ESM242"
output: 
  html_document:
    code_folding: hide
date: '2023-10-10'
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(nloptr)
library(knitr)
```


## Question 1: Mining Reserve {.tabset}


Suppose you own a mine with initial reserves $R_0$=1. You wish to determine the extraction rates, qt, that will maximize the sum of the present value of net benefits (PVNB) over a 10-year horizon t=0,1,…,9.  The net benefits from extraction are given by $\pi_t=c_0q_t+c_1q_t^2$ where $c_0$=0.8 and $c_1$=-0.1.  The discount rate is 0.05 and we’ll assume (at first) that any reserves left over in period 10 have a value of zero (i.e., $\lambda_0=0$). Find the extraction rates that maximize PVNB.  You will want to constrain extraction $q_t$ to be less than $R_t$ and for remaining reserves $R_{10}$ to be nonnegative.  

### A. 

What is the maximum PVNB with the initial set of parameters?

<span style="color: blue;">**Answer**</span>


```{r}
obj_func <- function(q, R_init, end_year, delta, c0, c1, lambda){
  
    R = vector(mode="numeric",length=(end_year+2)) ## make vector for reserves
    
    bens = vector(mode="numeric",length=0) ## make vector for benefits
    
    R[1] = R_init ## set initial reserve 
    bens[1] = c0*q[1] + c1*(q[1])^2 ## benefits are from the c0*q + c1*q^2
    
    for(i in 2:(end_year+1)){ ##for 2nd spot in vector to 10th (year 9)
      R[i] = R[i-1] - q[i-1] # every year R is diminished by the amount extracted
      bens[i] = c0*q[i] + c1*q[i]^2 ## benefits are given from the equation
      
    }
    
    R[end_year+2] = R[end_year+1] - q[end_year+1]
    bens[end_year+2] = R[end_year+2]*lambda
    
    t=seq(from=0,end_year+1) ##set amount of periods to add for present value 
  
    pv=((1/(1+delta))^t)*bens ##present value discounted
  
    npv=sum(pv) ## sum up present values, discounted
    
    return(-npv)
}




##make constraint function

R_constraint <- function(q, R_init = 1, end_year = 9, delta = 0.05, c0=0.8, c1=-0.1, lambda){
  const <- vector(mode="numeric", length=length(q))
  R <- vector(mode="numeric", length=length(q))

  
  R[1] = R_init
  
  for(i in 2:(end_year+1)){
    R[i] = R[i-1] - q[i-1]
    const[i] = q[i] - R[i]
  }
  return(const)
}



#Use these options with nloptr
options=list("algorithm"="NLOPT_LN_COBYLA",xtol_rel=1e-15,maxeval=16000)


 out=nloptr(x0=rep(0.1,times=10),eval_f=obj_func,
              eval_g_ineq = R_constraint,
              lb=rep(0,times=10),
              ub=rep(1,times=10),
              opts = options, 
              R_init = 1, end_year = 9, delta = 0.05, c0=0.8, c1=-0.1, lambda=0)

print(out$solution)
 tbl_1=round((out$solution),3)


kable(tbl_1, col.names = "Reserves Extracted")


```

The max PVNB is `r round(-out$objective,2)` 


### B.

How does the PVNB change when you assume that reserves left over in period 10 have a per unit value of $\lambda_{10}$=1.2? How do the optimal values of $q_t$ compare to those in part a? Explain.

<span style="color: blue;">**Answer**
</span>

```{r}


 out_1b=nloptr(x0=rep(0.1,times=10),eval_f=obj_func,
              eval_g_ineq = R_constraint,
              lb=rep(0,times=10),
              ub=rep(1,times=10),
              opts = options, 
              R_init = 1, end_year = 9, delta = 0.05, c0=0.8, c1=-0.1, lambda=1.2)

print(out_1b$solution)
print(-out_1b$objective)
 tbl_2=round((out_1b$solution),3)


kable(tbl_2, col.names = "Reserves Extracted")


```
The max PVNB is `r round(-out_1b$objective,2)`



### C.

What is the PVNB when $\lambda_{10}=1.4$ How do the optimal values of $q_t$ compare to those in part a? Explain.

<span style="color: blue;">**Answer**
</span>

```{r}


 out_1c=nloptr(x0=rep(0.1,times=10),eval_f=obj_func,
              eval_g_ineq = R_constraint,
              lb=rep(0,times=10),
              ub=rep(1,times=10),
              opts = options, 
              R_init = 1, end_year = 9, delta = 0.05, c0=0.8, c1=-0.1, lambda=1.4)

 tbl_3=round((out_1c$solution),3)
 
 

kable(tbl_3, col.names = "Reserves Extracted")

```

The max PVNB is `r round(-out_1c$objective,2)`



### D.

Even without using R, you could reason that the PVNB must be at least as large in parts b and c compared to part a. Explain.

<span style="color: blue;">**Answer**
</span>


If there is no benefit for receiving leftover reserves, then parts b and c would be the same as a. If there are benefits, optimizing the equation with lambda >1 would mean that parts b and c would have to yield a PVNB equal to part a or bigger. If nloptr optimization found that leaving extra unextracted reserves and receiving lamdba money for them (discounted) did NOT increase PVNB, then I would get the same answer in parts b and c as in part a. 


### E.

Now set $\lambda_{10}$ back to 0 and set $c_1$=0. Resolve and explain what happens to the optimal values of $q_t$ compare to those in part a. 

<span style="color: blue;">**Answer**
</span>

```{r}
out_1e=nloptr(x0=rep(0.1,times=10),eval_f=obj_func,
              eval_g_ineq = R_constraint,
              lb=rep(0,times=10),
              ub=rep(1,times=10),
              opts = options, 
              R_init = 1, end_year = 9, delta = 0.05, c0=0.8, c1=0, lambda=0)

 tbl_4=round((out_1e$solution),3)


kable(tbl_4, col.names = "Reserves Extracted")

```

The max PVNB is `r round(-out_1e$objective,2)`

If there are no consequences to extracting, there is no subtraction when calculating the present value of net benefits as time goes on, then you would extract all the reserves on the first go. Any after that would be discounted


