---
title: "HW 1"
author: "ESM242"
output: 
  html_document:
    code_folding: hide
date: '2023-10-10'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(nloptr)
library(knitr)
```


## Question 1: Mining Reserve {.tabset}


Suppose you own a mine with initial reserves $R_0$=1. You wish to determine the extraction rates, qt, that will maximize the sum of the present value of net benefits (PVNB) over a 10-year horizon t=0,1,…,9.  The net benefits from extraction are given by $\pi_t=c_0q_t+c_1q_t^2$ where $c_0$=0.8 and $c_1$=-0.1.  The discount rate is 0.05 and we’ll assume (at first) that any reserves left over in period 10 have a value of zero (i.e., $\lambda_0=0$). Find the extraction rates that maximize PVNB.  You will want to constrain extraction $q_t$ to be less than $R_t$ and for remaining reserves $R_{10}$ to be nonnegative.  

### A. 

What is the maximum PVNB with the initial set of parameters?

<span style="color: blue;">**Answer**</span>


```{r}
q1_eq <- function(q, t, delta, c0, c1){
  
      bens = ((1/(1+delta))^t)*((c0*q)+(c1*(q)^2))
      return(-bens)
}


#Use these options with nloptr
options=list("algorithm"="NLOPT_LN_COBYLA",xtol_rel=1e-15,maxeval=16000)

results = vector()
sum_bens = 0
total_ex = vector()
R=0.99 

for(i in 0:9){
  
 out = nloptr(x0=0,
          eval_f=q1_eq,
          lb=0,
          ub = R,
          opts = options,
          t=i, delta=0.05, c0=0.8, c1=-0.1)
 R = R - out$solution
}


 print(out$objective)
 print(out$solution)



```
**The maximum PVNB is `r sum_bens`**

### B.

How does the PVNB change when you assume that reserves left over in period 10 have a per unit value of $\lambda_{10}$=1.2? How do the optimal values of $q_t$ compare to those in part a? Explain.

<span style="color: blue;">**Answer**
</span>



### C.

What is the PVNB when $\lambda_{10}=1.4$ How do the optimal values of $q_t$ compare to those in part a? Explain.

<span style="color: blue;">**Answer**
</span>



### D.

Even without using R, you could reason that the PVNB must be at least as large in parts b and c compared to part a. Explain.

<span style="color: blue;">**Answer**
</span>



### E.

Now set $\lambda_{10}$ back to 0 and set $c_1$=0. Resolve and explain what happens to the optimal values of $q_t$ compare to those in part a. 

<span style="color: blue;">**Answer**
</span>



