---
title: "Forestry examples for ESM 242"
author: "Andrew Plantinga"
output: 
  html_document:
    code_folding: hide
date: '2022-06-18'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(nloptr)
library(knitr)
```



## Example 1: An optimal number of rotations {.tabset}


We seek to select the optimal harvest age T* for an even-aged forest to maximize the present discount value of profits from an infinite number of rotations.  The timber yield is given by $Q(T)=e^{a-b/T}$, where $a=10, b=53.27$, the cost of planting the stand is $c=300$, the price of a unit of timber is $p=1$, and the discount rate is $\delta=0.05$.  Thus, we choose $T^*$ to maximize $PNVB(T^*)=\frac{pQ(T^*)e^{-\delta T^*} - c}{1-e^{-\delta T^*}}$ where $T^* \gt 0$.


### A. 

What is the optimal rotation $T^*$?

<span style="color: blue;">**Solution**</span>


<span style="color: blue;">Build optimization problem through an objective function and constraint ensuring that harvest age $T^*$ is strictly positive</span>

```{r}


obj_fcn_tree1<-function(choice,a,b,c,p,discount){
  

  Q=exp(a-b/choice)
  benefits = (p*Q*exp(-discount*choice)-c)/(1-exp(-discount*choice))
  
  return(-benefits)
  
  
}


options=list("algorithm"="NLOPT_LN_COBYLA",xtol_rel=1e-8,maxeval=16000)



out_tree1=nloptr(x0=10,
          eval_f=obj_fcn_tree1,
          lb=0,
          opts = options,
          discount=0.03,a=10,b=53.27,c=300,p=1)



print ( out_tree1$solution )



```

If you increase the costs and hold all else constant (ex. C is now 600), you'll get a longer optimal rotation length because you'll want to delay the costs and have them be discounted more -- you want revenues closer in time and push costs off farther into the future but if the only thing you're changing is increasing the cost, it will increase the optimal rotation length.

If you increase the price (ex. p is now 2) -- the revenue will be greater and you'll want to bring that revenue closer in time and you will get a shorter rotation. 


If you discount the future LESS, aka are less impatient, you will have a longer optimal rotation period

**This model does not factor in the age of the trees**


## Example 2: Optimal single rotation with benefits from the standing stock

This problem has the same set up as example 1, but the stand will be grown for just one rotation and there will be positive benefits from the standing stock (e.g., wildlife habitat).  If the benefits in period $t$ from the standing stock are $B(t)=A(1-e^{-\alpha t})$, then the present value of cumulative benefits from time 0 to time $T$ are given by $PVB(T)=\frac{A}{\delta}[1-e^{-\delta T}]+\frac{A}{\alpha+\delta}[1-e^{-(\alpha+\delta)T}]$.  Assume $A=500$ and $\alpha=0.03$. The optimal rotation length $T^*$ is found with the code below.

```{r}


obj_fcn_tree2<-function(choice,a,b,c,p,discount,A,alpha){
  

  Q=exp(a-b/choice)
  timberbenefits = p*Q*exp(-discount*choice)-c
  stockbenefits = (A/discount)*(1-exp(-discount*choice))-(A/(alpha+discount))*(1-exp(-(alpha+discount)*choice))
  benefits=timberbenefits+stockbenefits  
    
    
  return(-benefits)
  
  
}


options=list("algorithm"="NLOPT_LN_COBYLA",xtol_rel=1e-8,maxeval=16000)



out_tree2=nloptr(x0=30,
          eval_f=obj_fcn_tree2,
          lb=0,
          opts = options,
          discount=0.05,a=10,b=53.27,c=300,p=1,A=500,alpha=0.03)



print ( out_tree2$solution )



```

If you're saying that the wildlife benefits are so high (e.g. protecting an endangered owl), then the time would be infinite when you try to solve because your solution would be that you should never cut down these forests.
