---
title: "ESM 242 HW 2"
author: "Andrew Plantinga"
date: '2022-07-01'
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(nloptr)
library(knitr)

```



## Question 1: Optimal fishery{.tabset}

This question examines the optimal management of a fishery. The stock dynamics are given by:

\begin{equation}
X_{t+1} = X_{t} + F(X_{t}) - Y_{t}
\end{equation}

where $F(X_{t})=rX_{t}(1-X_{t}/K)$ and $Y_t=qX_t E_t$.  Profits in each year are given by $\pi_t=pY_t-cE_t^2$.  Parameters values are given by:

\begin{equation}
\begin{aligned}
r& & 0.1\\
K& & 1 \\
q& & 0.01 \\
p& & 500 \\
c& & .01 \\
\delta& &0.05
\end{aligned}
\end{equation}

Assume the fishery manager chooses effort to maximize the present discounted value of profits over 50 periods, t=0,1,…,49.  Make sure to constrain the solution so that the amount of harvest $Y_{t}$ does not exceed the stock $X_{t}$ in any period.

### A.

What is the solution when the initial stock is 0.5?  Create a table that shows effort and stock in each period.  See fishery model in R_examples_1.Rmd for how to create a basic table from the nloptr output.

<span style="color: blue;">**Solution**</span>



```{r}

## Use this options list for nloptr

local_opts<-list("algorithm"="NLOPT_LN_COBYLA",xtol_rel=1e-15)
options=list("algorithm"="NLOPT_LN_AUGLAG",xtol_rel=1e-15,maxeval=16000,"local_opts"=local_opts)


```




### B.

What is the solution when the initial stock is 0.1?  Create a table that shows effort and stock in each period.

<span style="color: blue;">**Solution**</span>





### C.

What is the solution when the initial stock is 0.9?  Create a table that shows effort and stock in each period

<span style="color: blue;">**Solution**</span>




### D.

Compare the stock in the last period across the three solutions.  What do you notice?  Can you explain the pattern in terms of the fundamental equation of renewable resources?

<span style="color: blue;">**Solution**</span>


### E.

Compare the value of the objective function (i.e., the present value of profits) across the three solutions.  Explain the differences.  

<span style="color: blue;">**Solution**</span>
