---
title: "ESM 242 HW 2"
author: "Andrew Plantinga"
date: '2022-07-01'
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(nloptr)
library(knitr)

```



## Question 1: Optimal fishery{.tabset}

This question examines the optimal management of a fishery. The stock dynamics are given by:

\begin{equation}
X_{t+1} = X_{t} + F(X_{t}) - Y_{t}
\end{equation}

where $F(X_{t})=rX_{t}(1-X_{t}/K)$ and $Y_t=qX_t E_t$.  Profits in each year are given by $\pi_t=pY_t-cE_t^2$.  Parameters values are given by:

\begin{equation}
\begin{aligned}
r& & 0.1\\
K& & 1 \\
q& & 0.01 \\
p& & 500 \\
c& & .01 \\
\delta& &0.05
\end{aligned}
\end{equation}

Assume the fishery manager chooses effort to maximize the present discounted value of profits over 50 periods, t=0,1,…,49.  Make sure to constrain the solution so that the amount of harvest $Y_{t}$ does not exceed the stock $X_{t}$ in any period.

### A.

What is the solution when the initial stock is 0.5?  Create a table that shows effort and stock in each period.  See fishery model in R_examples_1.Rmd for how to create a basic table from the nloptr output.

<span style="color: blue;">**Solution**</span>



```{r}

opt_func <- function(E, stock_init, r, K, q, p, c, dsct, end_year){
  
  stock <- vector(mode = "numeric", length = (end_year+1))
  bens <- vector(mode = "numeric", length = (end_year+1))
  
  stock[1] = stock_init
  bens[1] = p*q*stock[1]*E[1] - c*((E[1])^2)
  
  for(i in 2:(end_year+1))
  {
    X = stock[i-1]
    stock[i] = X + r*X*(1-(X/K))-q*X*E[i-1]
    bens[i] <- p*q*stock[i]*E[i] - c*(E[i])^2
  }
  
   t=seq(from=0,end_year)
  
   pv=((1/(1+dsct))^t)*bens
   npv=sum(pv)
   return(-npv)
}
```


```{r}
constraint <- function(E, stock_init, r, K, q, p, c, dsct, end_year){
  
    stock <- vector(mode= "numeric", length=end_year+1)
    const <- vector(mode= "numeric", length=end_year+1)
    
    stock[1] = stock_init
  
  for(i in 2:(end_year+1)){
    X = stock[i-1]
    stock[i] = X + r*X*(1-(X/K))-q*X*E[i-1]
    const[i] = q*E[i]*(stock[i]) - stock[i]
  }
    return(const)
  
}
  
  
  
```

```{r,message=FALSE, warning=FALSE}

## Use this options list for nloptr

local_opts<-list("algorithm"="NLOPT_LN_COBYLA",xtol_rel=1e-15)
options=list("algorithm"="NLOPT_LN_AUGLAG",xtol_rel=1e-15,maxeval=16000,"local_opts"=local_opts)

 q1_out=nloptr(x0=rep(1,times=50),eval_f=opt_func,
              eval_g_ineq = constraint,
              lb=rep(0,times=50),
              opts = options, 
              stock_init = 0.5, r=0.1, K=1, q=0.01, p=500, c=0.01, dsct=0.05, end_year=49)


```
PVNB maximized is **$`r round(-q1_out$objective,2)`**


```{r}
  stock_calc <- function(E, X, r, K, q){
    current_stock = X+ r*X*(1-X/K) - q*X*E
    return(current_stock)
  }
```

```{r}
efforts_q1 <- q1_out$solution
 
 stocks_q1 <- vector(mode="numeric", length=length(efforts_q1))
 
 stocks_q1[1] = 0.5
 
 for (i in 2:length(efforts_q1)){
   stocks_q1[i] = stock_calc(efforts_q1[i-1], stocks_q1[i-1], r=0.1, K=1, q=0.01)
 }
  
results_q1 <- data.frame("Stocks" = round(stocks_q1,3), "Efforts"= round(efforts_q1,2)) 
kable(results_q1)
```



### B.

What is the solution when the initial stock is 0.1?  Create a table that shows effort and stock in each period.

<span style="color: blue;">**Solution**</span>


```{r}
 q2_out=nloptr(x0=rep(1,times=50),eval_f=opt_func,
              eval_g_ineq = constraint,
              lb=rep(0,times=50),
              opts = options, 
              stock_init = 0.1, r=0.1, K=1, q=0.01, p=500, c=0.01, dsct=0.05, end_year=49)
```

PVNB maximized is **$`r round(-q2_out$objective,2)`**

```{r}
efforts_q2 <- q2_out$solution
 
 stocks_q2 <- vector(mode="numeric", length=length(efforts_q2))
 
 stocks_q2[1] = 0.1
 
 for (i in 2:length(efforts_q2)){
   stocks_q2[i] = stock_calc(efforts_q2[i-1], stocks_q2[i-1], r=0.1, K=1, q=0.01)
 }
  
results_q2 <- data.frame("Stocks" = round(stocks_q2,3), "Efforts"= round(efforts_q2,2)) 
kable(results_q2)
```


### C.

What is the solution when the initial stock is 0.9?  Create a table that shows effort and stock in each period

<span style="color: blue;">**Solution**</span>


```{r}
 q3_out=nloptr(x0=rep(1,times=50),eval_f=opt_func,
              eval_g_ineq = constraint,
              lb=rep(0,times=50),
              opts = options, 
              stock_init = 0.9, r=0.1, K=1, q=0.01, p=500, c=0.01, dsct=0.05, end_year=49)
```

PVNB maximized is **$`r round(-q3_out$objective,2)`**

```{r}
efforts_q3 <- q3_out$solution
 
 stocks_q3 <- vector(mode="numeric", length=length(efforts_q3))
 
 stocks_q3[1] = 0.9
 
 for (i in 2:length(efforts_q3)){
   stocks_q3[i] = stock_calc(efforts_q3[i-1], stocks_q3[i-1], r=0.1, K=1, q=0.01)
 }
  
results_q3 <- data.frame("Stocks" = round(stocks_q3,3), "Efforts"= round(efforts_q3,2)) 
kable(results_q3)


```

### D.

Compare the stock in the last period across the three solutions.  What do you notice?  Can you explain the pattern in terms of the fundamental equation of renewable resources?

<span style="color: blue;">**Solution**</span>

With an initial stock of 0.5, the final stock is `r round(stocks_q1[50],3)`

With an initial stock of 0.1, the final stock is `r round(stocks_q2[50],3)`

With an initial stock of 0.9, the final stock is `r round(stocks_q3[50],3)`



### E.

Compare the value of the objective function (i.e., the present value of profits) across the three solutions.  Explain the differences.  

<span style="color: blue;">**Solution**</span>


With an initial stock of 0.5, the PVNB is $`r round(-q1_out$objective,2)`

With an initial stock of 0.1, the final stock is $`r round(-q2_out$objective,2)`

With an initial stock of 0.9, the final stock is $`r round(-q3_out$objective,2)`

It's going to be beneficial to have more stock at the beginning, because a effort goes a longer way closer to present. There can be high, relatively un-discounted values from catching with similar or smaller effort cost takeaways at the beginning